---
title: "1. Perceptions of the Behavior that Led to Anger"
author: "Pooya Razavi"
date: "2022-11-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#load libraries
package_list <- c("dplyr", "tidyr", "lmerTest")
lapply(package_list, require, character.only = TRUE)


#read in the dataset
df <- readxl::read_xlsx("preregistration_dataset.xlsx")

#set up categorical variables
df <- df %>% 
          mutate(NarrativeWritten = as.factor(df$NarrativeWritten),
                 NarrativeRelevant = as.factor(df$NarrativeRelevant),
                 Condition = as.factor(df$Condition))

#assigning values to factor levels
  levels(df$NarrativeWritten) <- c("No", "Yes")

  levels(df$NarrativeRelevant) <- c("No", "Yes", NA) 

  levels(df$Condition) <- c("justified", "nonjustified", NA)

#drop cases where condition = NA
  df1 <- df %>% filter(!is.na(Condition))

knitr::opts_chunk$set(echo = TRUE)
```


# Harm and threat

H(a) Justified anger events (compared to the unjustified ones) involve a stronger perception of harm and threat to the self and to others.

H(b) The difference between justified and unjustified anger in terms of perceptions of harm or threat is moderated by the target of harm, such that this difference is larger for harm to others (vs. harm to self). 

```{r}
#first examine the correlation between the 2 items for threat/harm to self and 2 items for threat/harm to other. If r > .5, proceed with the following analyses. If not, conduct these analyses for each item separately.

  cor.test(df1$harm_you, df1$threat_you) 
  cor.test(df1$harm_others, df1$threat_others) 
  
#assuming items are strongly correlated (r > .5)

#create the dataframe
harm_threat_df <- df1 %>% 
                      select(ResponseId, Condition, harm_you, harm_others, threat_you, threat_others) %>% 
                      mutate(harm.threat_self = ((harm_you + threat_you) / 2),
                             harm.threat_other = ((harm_others + threat_others) / 2))

#create a long dataframe
harm_threat_df_long <- harm_threat_df %>% 
                            tidyr::pivot_longer(cols = c(harm.threat_self, harm.threat_other),
                                                names_to = "target",
                                                names_prefix = "harm.threat_",
                                                values_to = "harm")


#the model: 
#the main effect of type of anger (i.e., condition) corresponds to H(a)
#the interaction effect corresponds to H(b)

harm_threat_df_long %>% 
  rstatix::anova_test(data = .,
                      dv = harm,
                      wid = ResponseId,
                      between = Condition,
                      within = target) %>% 
  rstatix::get_anova_table()



#difference between the two conditions in terms of harm to self
t.test(harm.threat_self ~ Condition, data = harm_threat_df)
effectsize::cohens_d(harm.threat_self ~ Condition, data = harm_threat_df)

#difference between the two conditions in terms of harm to others
t.test(harm.threat_other ~ Condition, data = harm_threat_df)
effectsize::cohens_d(harm.threat_other ~ Condition, data = harm_threat_df)
```


# Fairness

H: The target's behavior is perceived as less fair in the justified (vs. unjustified) anger condition.

```{r}
#first examine the correlation between the 2 items. If r > .5, proceed with the following analyses. If not, conduct these analyses for each item separately.

  cor.test(df1$beh_fair, df1$beh_justified) 

#assuming items are strongly correlated (r > .5) 
df1 <- df1 %>% 
          mutate(fair_just = ((beh_fair + beh_justified) / 2))

df1 %>% 
    t.test(fair_just ~ Condition, data = .)

df1 %>% 
    effectsize::cohens_d(fair_just ~ Condition, data = .)

```

# Norm violations

H(a): The target’s behavior during justified anger events (vs. unjustified ones) is judged as a stronger norm violation. 

H(b): The difference between justified and unjustified anger events is moderated by the type of norm violation, such that this difference is larger for injunctive norms (compared to descriptive norms). 

```{r}
#create the dataframe
norms_df <- df1 %>% 
                      select(ResponseId, Condition, how_acceptable, how_common) %>% 
                      mutate(injunctive = 6 - how_acceptable,
                             descriptive = 6 - how_common)
                      
                      

#create a long dataframe
norms_df_long <- norms_df %>% 
                            tidyr::pivot_longer(cols = c(injunctive, descriptive),
                                                names_to = "norm_type",
                                                values_to = "violation")


#the model: 
#the main effect of type of anger (i.e., condition) corresponds to H(a)
#the interaction effect corresponds to H(b)

norms_df_long %>% 
  rstatix::anova_test(data = .,
                      dv = violation,
                      wid = ResponseId,
                      between = Condition,
                      within = norm_type) %>% 
  rstatix::get_anova_table()



#difference between the two conditions in terms of injunctive norms
t.test(injunctive ~ Condition, data = norms_df)
effectsize::cohens_d(injunctive ~ Condition, data = norms_df)

#difference between the two conditions in terms of descriptive norms
t.test(descriptive ~ Condition, data = norms_df)
effectsize::cohens_d(descriptive ~ Condition, data = norms_df)

```

