---
title: "2. Perceptions of Target"
author: "Pooya Razavi"
date: "2022-11-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#load libraries
package_list <- c("dplyr", "tidyr", "lmerTest")
lapply(package_list, require, character.only = TRUE)


#read in the dataset
df <- readxl::read_xlsx("preregistration_dataset.xlsx")

#set up categorical variables
df <- df %>% 
          mutate(NarrativeWritten = as.factor(df$NarrativeWritten),
                 NarrativeRelevant = as.factor(df$NarrativeRelevant),
                 Condition = as.factor(df$Condition))

#assigning values to factor levels
  levels(df$NarrativeWritten) <- c("No", "Yes")

  levels(df$NarrativeRelevant) <- c("No", "Yes", NA) 

  levels(df$Condition) <- c("justified", "nonjustified", NA)

#drop cases where condition = NA
  df1 <- df %>% filter(!is.na(Condition))

knitr::opts_chunk$set(echo = TRUE)
```

# Causal Attribution

H: In anger events that are perceived as justified (compared to the unjustified ones), participants are more likely to consider the cause of anger to stem from the target’s internal and stable characteristics (as opposed to the external and changeable circumstances).

```{r}
#first, test the correlation between the two items: 
cor.test(df1$cause_circumst, df1$behave_same)

#If r > .5, create a composite and compare conditions:
    df1 <- df1 %>% 
              mutate(causal_attr = ((cause_circumst + behave_same) / 2))
    
    df1 %>% 
        t.test(causal_attr ~ Condition, data = .)
    
    df1 %>% 
        effectsize::cohens_d(causal_attr ~ Condition, data = .)

#If r < .5, compare conditions for each item:
  #cause: internal vs. external 
    df1 %>% 
        t.test(cause_circumst ~ Condition, data = .)
    
    df1 %>% 
        effectsize::cohens_d(cause_circumst ~ Condition, data = .)
  
  #cause: stable vs. the same 
    df1 %>% 
        t.test(behave_same ~ Condition, data = .)
    
    df1 %>% 
        effectsize::cohens_d(behave_same ~ Condition, data = .)
```

# Moral Character

H: In anger events perceived as justified (compared to the unjustified ones), the target is more likely to be seen as having weaker moral and ethical values. 

```{r}
#create the morality composite score using the items from Walker & Pitts (1998) 

  df_morality <- df1 %>% 
                    select(ResponseId, Condition, starts_with("wp_"), gw_honest, gw_principled)

  keys.list <- list(highly_moral=c("wp_concerned_right","wp_faithful","wp_faithful","wp_clear_values",
                                 "wp_lawabiding", "wp_strong_beliefs", "wp_distinguishes", "wp_dev_conscience",
                                 "wp_ethical", "gw_honest"))

  morality_scores <- psych::scoreItems(keys.list, df_morality)
  
  morality_scores
  
  df_morality <- cbind(df_morality, morality_scores$scores)


#compare conditions
    df_morality %>% 
        t.test(highly_moral ~ Condition, data = .)
    
    df_morality %>% 
        effectsize::cohens_d(highly_moral ~ Condition, data = .)


```

# Moral-Relational Character

Goodwin et al. (2014) demonstrate that judgments of morality and warmth are separable and can provide unique informational value for person perception. To examine the perceptions of the target along these two critical dimensions, I will first conduct dimensionality reduction analysis on the character judgments from Goodwin et al. (2014). 

```{r}
df_moral_relational <- df1 %>% 
                    select(ResponseId, Condition, starts_with("gw_"))

###Number of components will be decided based on scree plot, parallel analysis, and Velicor's MAP. If multiple options are suggested by these methods, I will conduct PCA for all options and make the decision about optimal factor solution based on component interpretability.

  #scree plot
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::scree(hline = -1)
  
  #parallel analysis
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::fa.parallel()
  
  #Velicor's MAP
  df_moral_relational %>% 
    select(starts_with("gw_")) %>% 
    psych::nfactors(n = 10)


  #PCA with n components (n will be replaced based on the outcome of analyses above)
  pca_n_component <- df_moral_relational %>% 
                        select(starts_with("gw_")) %>%  
                        psych::principal(., nfactors = n, rotate = "varimax") 
        
  n_comp_outcome <- psych::kaiser(pca_n_component, rotate = "Varimax") %>% psych::fa.sort()

  n_comp_outcome[["loadings"]] %>% 
                    knitr::kable(digits = 2) %>%
                    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>% 
                    kableExtra::kable_paper(full_width = F)

  #If there are cross-loading items (i.e., items that have high loadings on more than one component) in the PCA outcome, I will exclude such items one-by-one (starting with the ones that have the strongest cross-loading) until an interpretable solution without cross-loading is found. 

```

Based on the PCA results, I will create scale scores by averaging the corresponding items for each component. Next, I will compare the mean of emerging components between the two conditions.

```{r}

#hypothetical scenario: code for calculating scale scores for 2 components
  keys.list <- list(component1=c("...","...","...","..."),
                    component2=c("...","...","...","..."))

  moral_relation_scores <- psych::scoreItems(keys.list, df_moral_relational)
  
  moral_relation_scores
  
  df_moral_relational <- cbind(df_moral_relational, moral_relation_scores$scores)


#hypothetical scenario: compare conditions for each of the 2 components
  #component1
    df_moral_relational %>% 
        t.test(component1 ~ Condition, data = .)
    
    df_moral_relational %>% 
        effectsize::cohens_d(component1 ~ Condition, data = .)

  #component2
    df_moral_relational %>% 
        t.test(component2 ~ Condition, data = .)
    
    df_moral_relational %>% 
        effectsize::cohens_d(component2 ~ Condition, data = .)
```

