---
title: "04. Experiencer's Relational Processes"
author: "Pooya Razavi"
date: "2022-11-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#load libraries
package_list <- c("dplyr", "tidyr", "lmerTest")
lapply(package_list, require, character.only = TRUE)


#read in the dataset
df <- readxl::read_xlsx("preregistration_dataset.xlsx")

#set up categorical variables
df <- df %>% 
          mutate(NarrativeWritten = as.factor(df$NarrativeWritten),
                 NarrativeRelevant = as.factor(df$NarrativeRelevant),
                 Condition = as.factor(df$Condition))

#assigning values to factor levels
  levels(df$NarrativeWritten) <- c("No", "Yes")

  levels(df$NarrativeRelevant) <- c("No", "Yes", NA) 

  levels(df$Condition) <- c("justified", "nonjustified", NA)

#drop cases where condition = NA
  df1 <- df %>% filter(!is.na(Condition))
  
  
knitr::opts_chunk$set(echo = TRUE)
```

# Perception of Target's Regret

Research question: Do the perceptions of the target as regretful and apologetic vary depending on whether they consider the anger event as justified or unjustified? 

```{r}

  #Regret 
    df1 %>% 
        t.test(person_regret ~ Condition, data = .)
    
    df1 %>% 
        effectsize::cohens_d(person_regret ~ Condition, data = .)
  
  #Apology
    df1 %>% 
        t.test(person_apologize ~ Condition, data = .)
    
    df1 %>% 
        effectsize::cohens_d(person_apologize ~ Condition, data = .)

```

# Relationship Closeness

Hypothesis(a): Relationship closeness deteriorates following an anger event, both “short-term” (i.e., comparing closeness before and immediately after the event) and “long-term” (i.e., comparing closeness before the event and at the time of the study).

```{r}
closeness_df <- df1 %>% 
                      select(ResponseId, Condition, prior_closeness, immed_closeness, now_closeness)

#create a long dataframe
closeness_df_long <- closeness_df %>% 
                            tidyr::pivot_longer(cols = c(prior_closeness, immed_closeness, now_closeness),
                                                names_to = "time",
                                                values_to = "closeness") %>% 
                            mutate(time = as.factor(time)) %>% 
                            mutate(time = forcats::fct_relevel(time, "prior_closeness", "immed_closeness", "now_closeness"))
                      
  levels(closeness_df_long$time) <- c("prior", "immediate", "now")

#setup the contrasts
  contrasts(closeness_df_long$Condition) <- contr.sum
  contrasts(closeness_df_long$time) <- contr.sum  

#Overal model
  closeness_model <- lmer(closeness ~ time * Condition + (1 | ResponseId),
                          data = closeness_df_long)

  anova(closeness_model, type = "III") %>% knitr::kable()

#comparing closeness (a) before vs. immediately after the event 
  pairwise_mod <- emmeans::emmeans(closeness_model, pairwise ~ time)
  pairwise_mod[["contrasts"]][1]

#comparing closeness (b) before vs. now
  pairwise_mod[["contrasts"]][2]

```

Hypothesis(b): The long-term deterioration of relationship closeness is stronger for justified (vs. unjustified) anger experiences.

```{r}
longterm_mod <- emmeans::emmeans(closeness_model, pairwise ~ time | Condition)
  
  #justified: prior vs. now
  longterm_mod[["contrasts"]][2]
  #not justified: prior vs. now
  longterm_mod[["contrasts"]][5]


```

